---
title: "Lab 10 - Grading the professor"
author: "Sophie Boyd"
date: "3-6-26"
output: github_document
---

Here is a link to the [lab instructions](https://datascience4psych.github.io/DataScience4Psych/lab10.html).

## Load Packages and Data

```{r load-packages, message=FALSE, warning=FALSE}
library(tidyverse) 
library(tidymodels)
library(openintro)
```

# Part 1: Getting to know the outcome

## Exercise 1

```{r exercise1-plot, message=FALSE}
evals %>%
  ggplot(aes(x = score)) +
  geom_histogram(binwidth = .1) +
  labs(x = 'Score',
       y = 'Count',
       title = 'Distribution of Evaluation Scores')
```

```{r exercise1-summaries}
mean(evals$score)
median(evals$score)
min(evals$score)
max(evals$score)
```

Most students provided high evaluation ratings for their professors. The distribution of scores is skewed left, with most scores on the high end and a few scores on the low end.

## Exercise 2

```{r exercise2-scatter}
evals %>%
  ggplot(aes(x = bty_avg, y = score)) +
  geom_point() + 
  labs(x = 'Beauty Rating',
       y = 'Score',
       title = 'Evaluation Scores by Beauty Rating')
```

I don't see an especially clear trend in the relationship between beauty ratings and evaluation ratings on the scatterplot. There appears to be some organization into columns, where groups of professors with the same beauty ratings have a range of evaluation scores. 

## Exercise 3

```{r exercise3-jitter}
evals %>%
  ggplot(aes(x = bty_avg, y = score)) +
  geom_jitter() + 
  labs(x = 'Beauty Rating',
       y = 'Score',
       title = 'Evaluation Scores by Beauty Rating')
```

Jittering separates points that overlap each other on the scatterplot, providing more information about the density of the points. By jittering, we can see that scores are more concentrated at high values than low values across different beauty ratings, and especially among higher beauty ratings. (A positive association between beauty ratings and evaluation scores becomes clearer.) If someone were to only see the non-jittered plot, they might mistakenly assume that evaluation scores were evenly distributed among high and low values within groups of professors with the same beauty rating, when in fact higher evaluation ratings tended to be more common. 

# Part 2: Beauty as a predictor

## Exercise 1

```{r m_bty-model}
m_bty <- lm(score ~ bty_avg, data = evals)
summary(m_bty)
tidy(m_bty)
```

For each 1-point increase in beauty rating, the model predicts a .067-point increase in evaluation score. 

## Exercise 2

```{r m-bty-plot, message=FALSE}
evals %>%
  ggplot(aes(x = bty_avg, y = score)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = 'Beauty Rating',
       y = 'Evaluation Score')
```

## Exercise 3

- The slope is relatively small, indicating only a .067-point increase in evaluation score for each 1-point increase in beauty rating. 
- The predicted evaluation score for a professor with a beauty rating of 0 is 3.88. In context, I don't think this value is particularly meaningful. I believe that beauty ratings ranged from 1-10, in which case a rating of 0 would not be possible in our dataset. 
- The R-square value is .035, meaning that beauty ratings explain 3.5% of the variance in evaluation scores. Overall, it seems that a professor's attractiveness is not especially influential in student evaluations.
- At this stage, leaving the shading on the plot may lead someone to overstate how well the model fits the data. By looking at the line only, it is easier to where the model does not match up with the actual data.

# Part 3: Linear regression with a categorical predictor

##Exercise 1

```{r m_gen-model}
m_gen <- lm(score ~ gender, data = evals)
summary(m_gen)
```

The reference level for gender is female. The model predicts that, on average, male professors' evaluation ratings will be .14 points higher than female professors' evaluation ratings. 

Predicted evaluation score for female professor = 4.09 (intercept)
Predicted evaluation score for male professor = 4.23 (intercept + slope)

## Exercise 2

```{r rank-relevel}
evals <- evals %>%
  mutate(rank_relevel = relevel(rank, ref = "tenure track"))

evals <- evals %>%
  mutate(tenure_eligible = case_when(
    rank == "teaching" ~ "no",
    rank == "tenure track" ~ "yes",
    rank == "tenured" ~ "yes"
  ))
```

```{r rank-model}
m_rank <- lm(score ~ rank, data = evals)
summary(m_rank)
```

```{r rank-relevel-model}
m_rank_relevel <- lm(score ~ rank_relevel, data = evals)
summary(m_rank_relevel)
```

```{r tenure-eligible-model}
m_tenure <- lm(score ~ tenure_eligible, data = evals)
summary(m_tenure)
```

## Exercise 3

##### Model using original "rank" variable:
- Intercept: The average predicted evaluation score for teaching professors is 4.28.
- Slopes: On average, tenure track professors are expected to have evaluation scores .13 points lower than teaching professors. Tenured professors are expected to have evaluation scores .15 points lower than teaching professors. 

##### Model using "rank_relevel":
- Intercept: The average predicted evaluation score for tenure track professors is 4.15. 
- On average, teaching professors are expected to have evaluation scores .13 points higher than tenure track professors. Tenured professors are expected to have evaluation scores .02 points lower than tenure track professors.

##### Model using "tenure_eligible":
- Intercept: The average predicted evaluation score for teaching professors is 4.28.
- Slope: Compared to teaching professors, tenure eligible professors are predicted to have evaluation scores that are, on average, .14 points lower. 

Taken together, the three models communicate that teaching professors perform best on student evaluations. 

## Exercise 4

The R-square value is .012, meaning that rank only explains about 1.2% of the variance in evaluation scores. Overall, rank seems to play a small role. 

# Part 4: Multiple linear regression

```{r m_bty}
m_bty <- lm(score ~ bty_avg, data = evals)
summary(m_bty)
```

```{r m_bty_gen}
m_bty_gen <- lm(score ~ bty_avg + gender, data = evals)
summary(m_bty_gen)
```

## Exercise 1

The coefficient for beauty increased slightly and the associated p-value decreased after adding gender as a predictor.

## Exercise 2

Yes, gender is an individual significant predictor of evaluation score. 

## Exercise 3

I would say that adding gender as a predictor was somewhat helpful. The model with beauty alone explained 3.3% of the variance in evaluation scores, whereas the model that added gender explained 5.5% of the variance in evaluation scores. 

## Exercise 4

```{r m_bty_rank}
m_bty_rank <- lm(score ~ bty_avg + rank, data = evals)
summary(m_bty_rank)
```

- Holding rank constant, a one-unit increase in beauty rating predicts a .07-point increase in evaluation score.

- Holding beauty rating constant, tenure-track professors are predicted to have evaluation scores .016 points lower than teaching professors. 

# Part 5: The search for the best model

## Exercise 1

I would expect the number of professors teaching sections of a course to be the weakest predictor of evaluation scores because students are typically only aware of/focused on the professor who taught their section when completing evaluations.

## Exercise 2

```{r m_cls_profs}
m_cls_profs <- lm(score ~ cls_profs, data = evals)
summary(m_cls_profs)
```

## Exercise 3

If the model already included cls_perc_eval and cls_students, it would be redundant to include cls_did_eval, as the number of students who completed the evaluation does not communicate any new information beyond the total number of students in the class and the percentage of students who completed the evaluation. It would not explain additional variance in evaluation scores.


## Exercise 4

```{r full-model}
m_full <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval + cls_students + cls_level + cls_profs + cls_credits + bty_avg, data = evals)
summary(m_full)
```

## Exercise 5

```{r best-model}
m_best <- lm(score ~ ethnicity + gender + language + age + cls_perc_eval + cls_credits + bty_avg, data = evals)
summary(m_best)
```

Predicted eval score = 3.45 + .20(ethnicity) + .18(gender) - .16(language) - .01(age) + .01(cls_perc_eval) + .52(cls_credits) + .06(bty_avg)

## Exercise 6

- beauty rating (numerical): Holding other predictors constant, each one-unit increase in beauty rating predicts a .06-point increase in evaluation score.

- class credits (categorical): Holding other predictors constant, professors of one-credit classes are predicted to receive evaluation scores .52 points higher than professors of multicredit classes. 

## Exercise 7

A professor at UT Austin with a high evaluation score would be a young, white, male, English-speaking professor teaching a one-credit class with a high percentage of students who completed the course evaluation and a high beauty rating. 

## Exercise 8

I would not be entirely comfortable with generalizing these results to other universities, the makeup of the student body varies widely across different universities, which would likely influence how students evaluate their professors. For example, given that UT Austin is a large state university, the results might not apply to a smaller liberal arts school. 
